Consistency scores in text data

### 2020-07-24 outputs/documents/paper.Rmd
Write-up minor revisions

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: pdflatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
biblio-style: apalike
title: "Consistency scores in text data"
thanks: "Thank you to X, Y and Z for helpful comments. A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/. An R package for larger scale applications is available: **HERE**. Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Rohan Alexander
  affiliation: University of Toronto
- name: Ke-Li Chiu
  affiliation: University of Toronto  
abstract: "In this paper we introduce a process to clean the text extracted from PDFs using various methods from natural language processing. Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus. To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model. This is used to monitor changes during the cleaning process, and to compare the messiness of different texts. The methods that we consider are: n-grams, SOMETHING, GPT-2, and GPT-3. We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
keywords: "text-as-data; natural language processing; GPT-3; quantitative analysis."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# spacing: double
endnote: no
graphics: yes
---

@@

We consider various natural language processing methods. These include n-grams, **SOMETHING**, as well as GPT2 and GPT-3. The n-gram approach involves identifying two, three, or more, words that are commonly found together. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance an equally reasonable two-gram involving the word 'good' is 'good work'. For that reason we consider pre-trained generative models. These include **SOMETHING**, which works **SOMEHOW**. We also consider GPT-2 and GPT-3 from OpenAI. GPT-2 and GPT-3 are pre-trained generative unsupervised language models. They differ in the number of parameters. Generative Pretrained Transformer (GPT)-2 has *X* parameters and GPT-3 has more than 175 billions parameters. While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot.

@@

One issue in this process is that OCR is not perfect. Hence there is a need to correct the output before it can be analysed. Traditionally, word-correction techniques evaluate errors one by one without considering the context of the surrounding words. This was no longer the case in modern correction techniques as statistical language models (SLMs) and feature-based methods have been used for context-sensitive correction (citation). Without exception, all human languages have some words that co-occur more frequently with others. Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (citation). In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$. This conditional probability is the foundation of an n-gram language model.

@@

"We" is a unigram, "We had" is a bigram, "We had been" is a trigram, and "We had been wandering" is a 4-gram. On the other hand, the trigrams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language. Moreover, N-gram then enables us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next. Consider the two sentences: "We had been wandering," and "We had been wangling." The former is likely to be more frequently encountered in a training corpus. Thus, the n-gram would assign a higher probability to "wandering" than to "wangling."

To predict a word appearing next, we have to take the sequence of preceeding words into account, which requires knowing the probability of the sequence of words. The probability of a sequence appearing in a corpus follows the chain rules:

@@

However, the likelihood that more and more words will occur next to each other in an indentical sequence becomes smaller and smaller, making the prediction difficult. Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words **(Keli add citation)**. More specifically

@@

Because the bigram model looks into one word in the past, the n-gram under Markov assumption can then be generalized into a bigram model with n being any number. The general equation is

P($w_{n} | w^{n-1}_{1}) \approx$ P($w_{n} | w^{n-1}_{n - N+1})$)

The application of n-gram is versatile. N-gram is widely applied in text prediction, spelling-correction, machine translation. In our application, we use a trigram-based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability. [add more about how we use n-gram once the functions in the app are in place]

@@

### Article: Unsupervised Learning of Sentence Embeddings using Compositional n-Gram Features (2017)
Embedding of sentences (sequence of words with semantic representations) compositional n-gram features.

@@

In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing **(Ke-Li add ref)**. Pre-trained models remove the need for researchers to create training datasets or devote computational resources to this stage. Because of this, there has been substantial development in NLP research.

OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was announced by OpenAI in June 2020 **(Ke-Li add ref)**. OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**. GPT-3 is the third Generative Pre-trained Transformer language model publicly released by OpenAI. It has '175 billion parameters' and can learn NLP tasks with just a few examples. OpenAI claims that the text generation by GPT-3 is 'difficult for humans to distinguish from authentic human content' **(Ke-Li add ref)**. Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited. On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years. Alt, Hubner and Henning extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text. They had shown a milestone in predicting larger range of distinct relation types with higher confidence **(Ke-Li add ref)**. In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning **(Ke-Li add ref)**. Bird, Faria, Ekart, Premebida and Ayrosa employed GPT-2 as part of the solution to generate augmented data for classification. GPT-2 also helped NLP researchers tackling the task of textual paraphrasing **(Ke-Li add ref)**. Hedge and Patil used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.  
**(Ke-Li please find papers that use it, describe them, and add refs - look on arXiv)**.

GPT-3 is a Transformer-based model that has over 175 billions of parameters. **(Keli add citation)**. Proposed by Vaswani et al., the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency **(Keli add citation)**. The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**. In the past few years, the growth of parameters of transformer language models expands rapidly. The first generation GPT has 110 million parameters. Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor . The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**. GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning. Few-shot learning means that the model  **(Keli add citation)**. 
As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task. **(Ke-Li add more detail about what it needs and what it outputs and how it works).**. For example, OpenAI team trained GPT-3 to generate "news articles". The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself. Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets. The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories **(Keli add citation)**. This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields. **(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**

>>>>

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: pdflatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
biblio-style: apalike
title: "Consistency scores in text data"
thanks: "Thank you to X, Y and Z for helpful comments. A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/. An R package for larger scale applications is available: **HERE**. Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Rohan Alexander
  affiliation: University of Toronto
- name: Ke-Li Chiu
  affiliation: University of Toronto  
abstract: "In this paper we introduce a process to clean the text extracted from PDFs using various methods from natural language processing. Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus. To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model. This is used to monitor changes during the cleaning process, and to compare the messiness of different texts. The methods that we consider are: n-grams, [Google’s Word2vec, Standford's GloVe], GPT-2, and GPT-3. We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
keywords: "text-as-data; natural language processing; GPT-3; quantitative analysis."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# spacing: double
endnote: no
graphics: yes
---

@@

We consider various natural language processing methods. These include n-grams, [Google’s Word2vec, Standford's GloVe], as well as GPT2 and GPT-3. The n-gram approach involves identifying two, three, or more, words that are commonly found together. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance an equally reasonable two-gram involving the word 'good' is 'good work'. For that reason we consider pre-trained word embedding models. These include [Google’s Word2vec, Standford's GloVe], which works [representing words as dense vectors that capture their semantic relationships with each other]**SOMEHOW**. We also consider pre-trained generative models such as GPT-2 and GPT-3 from OpenAI. GPT-2 and GPT-3 are pre-trained generative unsupervised language models. They differ in the number of parameters. Generative Pretrained Transformer (GPT)-2 has *X* parameters and GPT-3 has more than 175 billions parameters. While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot.

@@

One issue in this process is that OCR is not perfect. Hence there is a need to correct the output before it can be analysed. Traditionally, word-correction techniques evaluate errors one by one without considering the context of the surrounding words. This was no longer the case in modern correction techniques as statistical language models (SLMs) and feature-based methods have been used for context-sensitive correction (citation). Without exception, all human languages have some words that co-occur more frequently with others. Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (@brown1992class). In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$. This conditional probability is the foundation of an n-gram language model.

@@

"We" is a uni-gram, "We had" is a bi-gram, "We had been" is a tri-gram, and "We had been wandering" is a 4-gram. On the other hand, the tri-grams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language. Moreover, using n-gram models then enable us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next. Consider the two sentences: "We had been wandering," and "We had been wangling." The former is likely to be more frequently encountered in a training corpus. Thus, the n-gram would assign a higher probability to "wandering" than to "wangling."

To predict which word appears next, we have to take the sequence of preceeding words into account, which requires knowing the probability of the sequence of words. The probability of a sequence appearing in a corpus follows the chain rule:

@@

However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making the prediction difficult. Alternatively, we can approximate the probability of a word depending on only the previous word; this is known as Markov assumption. Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words (@brown1992class). More specifically:

@@

Because the bi-gram model looks only the immediate preceding word, an n-gram model under Markov assumption can then be reduced into a bi-gram model with n being any number. The general equation is:

> P($w_{n} | w^{n-1}_{1}) \approx$ P($w_{n} | w^{n-1}_{n - N+1})$)

The application of n-gram is versatile. N-gram models are widely applied in text prediction, spelling-correction, machine translation **(Keli add citation)**. In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability. [add more about how we use n-gram once the functions in the app are in place]

@@

@@

In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing **(Ke-Li add ref)**. Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training. Because of this, there has been substantial development in NLP research.

OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was announced by OpenAI in June 2020 **(Ke-Li add ref)**. OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**. GPT-3 is the third Generative Pre-trained Transformer language model publicly released by OpenAI. It has '175 billion parameters' and can learn NLP tasks with just a few examples. OpenAI claims that the text generation by GPT-3 is 'difficult for humans to distinguish from authentic human content' (@brown2020language). Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited. On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years. @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text. They had shown a milestone in predicting larger range of distinct relation types with higher confidence **(Ke-Li add ref)**. In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning **(Ke-Li add ref)**. @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification. GPT-2 also helped NLP researchers tackling the task of textual paraphrasing **(Ke-Li add ref)**. @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.  
**(Ke-Li please find papers that use it, describe them, and add refs - look on arXiv)**.

GPT-3 is a Transformer-based model that has over 175 billions of parameters (@brown2020language). Proposed by @vaswani2017attention, the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency **(Keli add citation)**. The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**. In the past few years, the growth of parameters of transformer language models expands rapidly. The first generation GPT has 110 million parameters. Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor . The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**. GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning. Few-shot learning means that the model  **(Keli add citation)**. 
As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task. **(Ke-Li add more detail about what it needs and what it outputs and how it works).**. For example, OpenAI team trained GPT-3 to generate "news articles". The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself. Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets. The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories (@brown2020language). This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields. **(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**
